---
title: "jsonvalidate"
author: "Rich FitzJohn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{jsonvalidate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

``` {r echo=FALSE, results="hide"}
knitr::opts_chunk$set(error=FALSE)
```

This package wraps
[is-my-json-valid](https://github.com/mafintosh/is-my-json-valid)
using [V8](https://cran.r-project.org/package=V8) to do JSON schema
validation in R.

You need a JSON schema file; see
[json-schema.org](http://json-schema.org) for details on writing
these.  Often someone else has done the hard work of writing one
for you, and you can just check that the JSON you are producing or
consuming conforms to the schema.

The examples below come from the [JSON schema
website](http://json-schema.org/example1.html)

They describe a a JSON based product catalog, where a product which
has an id, a name, a price, and an optional set of tags.  A JSON
represenation of a product is:

```json
{
    "id": 1,
    "name": "A green door",
    "price": 12.50,
    "tags": ["home", "green"]
}
```

The schema that they derive looks like this:

```json
{
    "$schema": "http://json-schema.org/draft-04/schema#",
    "title": "Product",
    "description": "A product from Acme's catalog",
    "type": "object",
    "properties": {
        "id": {
            "description": "The unique identifier for a product",
            "type": "integer"
        },
        "name": {
            "description": "Name of the product",
            "type": "string"
        },
        "price": {
            "type": "number",
            "minimum": 0,
            "exclusiveMinimum": true
        },
        "tags": {
            "type": "array",
            "items": {
                "type": "string"
            },
            "minItems": 1,
            "uniqueItems": true
        }
    },
    "required": ["id", "name", "price"]
}
```

This ensures the types of all fields, enforces presence of `id`,
`name` and `price`, checks that the price is not negative and
checks that if present `tags` is a unique list of strings.

There are two ways of passing the schema in to R; as a string or as
a filename.  If you have a large schema loading as a file will
generally be easiest!  Here's a string representing the schema
(watch out for escaping quotes):
``` {r }
schema <- '{
    "$schema": "http://json-schema.org/draft-04/schema#",
    "title": "Product",
    "description": "A product from Acme\'s catalog",
    "type": "object",
    "properties": {
        "id": {
            "description": "The unique identifier for a product",
            "type": "integer"
        },
        "name": {
            "description": "Name of the product",
            "type": "string"
        },
        "price": {
            "type": "number",
            "minimum": 0,
            "exclusiveMinimum": true
        },
        "tags": {
            "type": "array",
            "items": {
                "type": "string"
            },
            "minItems": 1,
            "uniqueItems": true
        }
    },
    "required": ["id", "name", "price"]
}'
```

Create a validator:
``` {r }
v <- jsonvalidate::json_validator(schema)
```

If we'd saved he json to a filename this would work too:
``` {r }
writeLines(schema, "schema.json")
v <- jsonvalidate::json_validator("schema.json")
```

The returned object is a function that takes as its first argument
a json string, or a filename of a json file.  The empty list will
fail validation because it does not contain any of the required fields:
``` {r }
v("{}")
```

To get more information on why the validation fails, add `verbose=TRUE`:
``` {r }
v("{}", verbose=TRUE)
```

The attribute "errors" is a data.frame and is present only when the
json fails validation.  The error messages come straight from
`is-my-json-valid` and they may not always be that informative.

Alternatively, to throw an error if the json does not validate, add
`error=TRUE` to the call:
``` {r error=TRUE}
v("{}", error=TRUE)
```

And to continue validating after the first error, pass `greedy=TRUE`:
``` {r }
v("{}", verbose=TRUE, greedy=TRUE)
```

which will sometimes show more errors.

The JSON from the opening example works:
``` {r }
v('{
    "id": 1,
    "name": "A green door",
    "price": 12.50,
    "tags": ["home", "green"]
}')
```

But if we tried to enter a negative price it would fail:
``` {r }
v('{
    "id": 1,
    "name": "A green door",
    "price": -1,
    "tags": ["home", "green"]
}', verbose=TRUE)
```

...or duplicate tags:
``` {r }
v('{
    "id": 1,
    "name": "A green door",
    "price": 12.50,
    "tags": ["home", "home"]
}', verbose=TRUE)
```

or just basically everything wrong:
``` {r }
v('{
    "id": "identifier",
    "name": 1,
    "price": -1,
    "tags": ["home", "home", 1]
}', verbose=TRUE)
```

The `data.tags.2` name comes from within the `is-my-json-valid`
source, and may be annoying to work with programmatically.

There is also a simple interface where you take the schema and the
json at the same time:
``` {r }
json <- '{
    "id": 1,
    "name": "A green door",
    "price": 12.50,
    "tags": ["home", "green"]
}'
jsonvalidate::json_validate(json, schema)
```

``` {r echo=FALSE, results="hide"}
file.remove("schema.json")
```
